---
title: "Uber driver survey November 2018"
author: "Michelle Hoogenhout"
date: "14 November 2018"
output: html_document
---

# Purpose 
Design a training pilot

# Hypotheses in designing the pilot:

UBER drivers...
- want higher-value economic opportunities / to learn skills
- are interested in digital careers: Trades, digital, business
- have the aptitude for high-value digital careers
- are prepared to study part-time alongside UBER driving long enough to develop employable skills
- are prepared to switch to being an employee or freelancer rather than UBER driver "entrepreneur" on completion of the programme

```{r setup, include=FALSE}

library(tidyverse) #data wrangling
library(magrittr) #pipe
library(ggplot2) #plot
library(scales) #for plotting
library(tm) #word count 
library(tau)

source("uber_getdatasets.R")

```

```{r functions, include = F}

#get percentage of TRUE observations
getpercentage <- function(var){
 x <- length(which(var == 1))/length(var)*100
 print(round(x,0))
}

#basic bargraph
bargraph <- function(data, x, title = ""){
 p <- ggplot(data, aes(x = x))
 p + geom_bar(aes(y = (..count..)/sum(..count..))) + 
  scale_y_continuous(labels=percent) + 
  labs(x = "", y = "") +
  ggtitle(title) +
  theme(axis.text.x = element_text(angle = 40, size = 16, hjust = 1),
        axis.text.y = element_text(size = 16))
 
}

#export graph
exportgraph <- function(graphname, exportname, width=8, height=6){
 png(filename = exportname, 
     type="cairo",
     units="in", 
     width=width, 
     height=height, 
     pointsize=12, 
     res=96)
 return(graphname)
 dev.off()
 
}

#count number of observations fitting within defined category in a list
countcats <- function(category, list){
 length(grep(paste(category,collapse="|"), 
             list, value=TRUE))
}

```


```{r get data}

#change according to dataset required

DATA <- alldata
WILLSTUDY <- willstudy_all

#only for alldata
DATA$sample <- ifelse(DATA$id %in% c(1:999), "GLH", "expo")
                 
```


## Drivers want higher-value economic opportunities / to learn skills

What percentage of drivers want to study?
Mode and distibution of willingness to spend
Count of certificate, new skill, job

```{r cost outcome willing}

#how many drivers are willing to study?
WillingStudyGraph <- bargraph(DATA, DATA$WillingnessStudy, "Willing to Study Part-Time?")

costlevels <- c("Nothing", "Less than R200", "R200 - R499", 
                "R500 - R999", "R1000 - R1999", "R2000 - R2999")

#there is one NA value from someone who otherwise completed all questiosn - drop NA
willstudy_complete <- WILLSTUDY[which(!is.na(WILLSTUDY$PreferredCost)),]

costgraph <- bargraph(willstudy_complete, willstudy_complete$PreferredCost, "Preferred Cost") +
 scale_x_discrete(labels = costlevels)

outcomegraph <- bargraph(WILLSTUDY, WILLSTUDY$DesiredOutcome, "Desired Outcome")

```


## What skills do drivers want to learn?
rank skills offered; group by trade, business, digital, other

```{r skill types}

#List all Skills in vector
temp <- unlist(strsplit(as.character(WILLSTUDY$Skills), c(", ", "; ", " AND ")))
skillslist <- sort(trimws(unlist(strsplit(temp, ",")), which = "both"))

#categories
source("skillcodes.R")    
           
skillcats <- list(TRADESKILLS, BUSINESS_SKILLS, IT_SKILLS, MARKETING, 
                  ENGINEERING, HEALTH, CREATIVE, ENGLISH, AGRICULTURE)



#count skills by category
skillscount <- list()

for (cat in seq_along(skillcats)){
 skillscount[[cat]] <- countcats(skillcats[[cat]], skillslist)
}

#Bind to df
count <- unlist(skillscount)
skill_df <- data.frame(c("Trade", "Business", "IT", "Marketing", "Engineering", 
                         "Health", "Creative", "English", "Agriculture"), count)

colnames(skill_df) <- c("Category", "Count")

#arrange largest to smallest
skill_df$Category <- factor(skill_df$Category, 
                            levels = skill_df$Category[order(-skill_df$Count)])

#graph popular skill areas 
SkillGraph <- ggplot(skill_df, aes(Category, Count)) +
 geom_bar(stat = "identity") + 
 ggtitle("Popular skill areas") +
 theme(axis.text.x = element_text(angle = 40, size = 16, hjust = 1),
       axis.text.y = element_text(size = 16)) +
 labs(x = "", y = "Count")

```


## Drivers are prepared to study for long enough to develop skills

```{r willingness to learn, echo = F}

LengthStudyGraph <- bargraph(WILLSTUDY, WILLSTUDY$LengthWillingStudy, "Length Willing to Study")
TimeCommitmentGraph <- bargraph(WILLSTUDY, WILLSTUDY$DailyTimeCommitment, "Daily Time Commitment")
AttendanceGraph <- bargraph(WILLSTUDY, WILLSTUDY$AttendanceFrequency, "Frequency of Attendance")

```


## Drivers have the aptitude to learn new skills


```{r study completion rates, echo = F}

studiedprev <- WILLSTUDY %>% filter(StudiedPreviously == "Yes")

#TODO: arrange largest to smallest?

#Provide shortened names for graph
barriers = c("Education support", "Money", "Time", "Started Working", "Prefer not to say", 
             "Personal", "Motivation", "Too difficult", "Unaccredited Institution")

StudiedGraph <- bargraph(WILLSTUDY, WILLSTUDY$StudiedPreviously, "Studied Previously?")

CompletedStudyGraph <- bargraph(studiedprev, studiedprev$CompletedStudies,"Completed Studies?")

#filter out those who are currently studying from incomplete list
incomplete_study <- studiedprev %>% 
 filter(CompletedStudies == "No" & ReasonNotCompleting != "Currently Studying")
 
NotCompletedStudyGraph <- bargraph(incomplete_study, incomplete_study$ReasonNotCompleting, "Barriers to Completion") +
 scale_x_discrete(labels = barriers)

```


## Drivers are prepared to switch to employee/ freelancer role

```{r switch position}

SwitchPositionGraph <- bargraph(WILLSTUDY, WILLSTUDY$SwitchPosition, "Prepared to Switch Position")

```


## Drivers have access to computers and internet for self study

Of the drivers surveyed, `r getpercentage(DATA$laptop)` have laptops, `r getpercentage(DATA$HomeInternet)` have internet at home. Most drivers access the internet through their mobile phone. 


```{r laptop internet}

#what percentage of drivers have access to a laptop/pc and internet?

getpercentage(DATA$laptop)
getpercentage(DATA$HomeInternet)
getpercentage(DATA$FriendsPC)


InternetGraph <- bargraph(WILLSTUDY, WILLSTUDY$InternetAccess, "Method of Internet Access")

```

```{r graph export}

graphlist <- list(c(outcomegraph, "figs/Outcome.png"),
                  c(costgraph, "figs/Cost.png"),
                  c(LengthStudyGraph, "figs/LengthStudy.png"),
                  c(LengthStudyGraph, "figs/LengthStudy.png"),
                  c(TimeCommitmentGraph, "figs/TimeCommitment.png"),
                  c(AttendanceGraph, "figs/Attendance.png"),
                  c(StudiedGraph, "figs/StudiedBefore.png"),
                  c(CompletedStudyGraph, "figs/CompletedStudies.png"),
                  c(NotCompletedStudyGraph, "figs/ReasonsNotCompletedStudy.png"),
                  c(SwitchPositionGraph, "figs/SwitchPosition.png"),
                  c(InternetGraph, "figs/InternetAccess.png"))
# 
# for (graph in seq_along(graphlist)){
#   exportgraph(bquote(.(graphlist[graph][1])), bquote(.(graphlist[graph[2]])))
# }

exportgraph(WillingStudyGraph, "figs/alldata/WillingToStudy.png")
exportgraph(outcomegraph, "figs/alldata/Outcome.png")
exportgraph(costgraph, "figs/alldata/Cost.png")
exportgraph(LengthStudyGraph, "figs/alldata/LengthStudy.png")
exportgraph(TimeCommitmentGraph, "figs/alldata/TimeCommitment.png")
exportgraph(AttendanceGraph, "figs/alldata/Attendance.png")
exportgraph(StudiedGraph, "figs/alldata/StudiedBefore.png")
exportgraph(CompletedStudyGraph, "figs/alldata/CompletedStudies.png")
exportgraph(NotCompletedStudyGraph, "figs/alldata/ReasonsNotCompletedStudy.png")
exportgraph(SwitchPositionGraph, "figs/alldata/SwitchPosition.png")
exportgraph(InternetGraph, "figs/alldata/InternetAccess.png")
exportgraph(SkillGraph, "figs/alldata/SkillCategories.png")


dev.off()
```

