---
title: "Uber driver survey November 2018"
author: "Michelle Hoogenhout"
date: "14 November 2018"
output: html_document
---

# Purpose 
Design a training pilot

#Hypotheses in designing the pilot:

UBER drivers...
- want higher-value economic opportunities / to learn skills
- are interested in digital careers: Trades, digital, business
- have the aptitude for high-value digital careers
- are prepared to study part-time alongside UBER driving long enough to develop employable skills
- are prepared to switch to being an employee or freelancer rather than UBER driver "entrepreneur" on completion of the programme

```{r setup, include=FALSE}

library(ggplot2)
library(tidyverse)
library(magrittr)
library(psych)
library(cowplot) #for plotting
library(scales) #for plotting
library(readxl) #for excel files

```

**want higher-value economic opportunities / to learn skills**

What percentage of drivers want to study?
Mode and distibution of willingness to spend
Count of certificate, new skill, job

```{r define factor levels}

#define order of factor levels
OrderLengthStudy<- c("Less than 6 months", "6 months - 1 year", "1 - 2 years",
                    "3 - 5 years", "More than 5 years")

OrderTimeCommitment<- c("1 - 2 hours", "3 -5 hours", "6 - 10 hours",
                   "10 - 15 hours", "16 - 20 hours", "More than 20 hours")

OrderFreqAttendance<- c("Every weekday", "Twice a week", "Once a week", "Once a month")


#combine all variance of currntly studying
studying_list = c("Currently busy with studying", 
                  "currently studying", 
                  "Currently studying", 
                  "Studies in progress",
                   "I am currently busy studying at college")

```

```{r get data, include=FALSE}

data <- readxl::read_xlsx("Umuzi Driver Education Revised.xlsx")

#rename and create factors, defining order and collapsing levels of currently studying
data %<>% rename(id = "Response ID", 
                 WillingnessStudy = "Are you willing to study part-time while operating on Uber?",
                 LengthWillingStudy = "For how long would you be willing to study while operating on Uber?",
                 Skills = "Which skills would you most like to learn? Select up to three skills.",
                 DesiredOutcome = "What is most important to you at the end of studying?",
                 WeeklyTimeCommitment = "How much time can you commit on a weekly basis to studying?",
                 AttendanceFrequency = "How often would you be able to attend classes, in-person, at a local campus in your city?",
                 
                PreferredClassFormat = "What format would you prefer for the classes?",
                ResourceAccess = "I have access to",
                InternetAccess = "How do you access the internet when you not operating on Uber?",
                PreferredCost = "How much would you be willing to pay monthly for your part-time studying?",
                SwitchPosition = "If you were given the opportunity to start a new career, would you be willing to start in a junior position?",
                StudiedPreviously = "Have you studied before?",
                StudySubject = "If \" yes\", what did you study?",
                CompletedStudies = "Did you complete your studies?",
                ReasonNotCompleting = "What prevented you from completing your studies?",
                FocusGroup = "Would you be willing to attend a focus group to talk about your goals for further education and skills?") %>% 
  mutate_all(funs(factor)) %>%
  mutate(LengthWillingStudy = factor(LengthWillingStudy, levels = OrderLengthStudy),
         WeeklyTimeCommitment = factor(WeeklyTimeCommitment, levels = OrderTimeCommitment),
         AttendanceFrequency = factor(AttendanceFrequency, levels = OrderFreqAttendance),
         ReasonNotCompleting = fct_collapse(ReasonNotCompleting, `Currently Studying` = studying_list))

#change Studied Previously frm yes to Currently Studying
levels(data$StudiedPreviously) <- c("No", "Yes", "Currently Studying")
data[which(data$ReasonNotCompleting == "Currently Studying"),]$StudiedPreviously <- "Currently Studying"

```


**What skills do drivers want to learn?**
rank skills offered; group by trade, business, digital, other

```{r}


```

**Drivers are prepared to study for long enough to develop skills**

```{r willingness to learn, echo = F}

willstudy <- data[which(data$WillingnessStudy == "Yes"),]
 
bargraph <- function(data, x, xtitle){
  p <- ggplot(data, aes(x = x))
  p + geom_bar(aes(y = (..count..)/sum(..count..))) + 
    scale_y_continuous(labels=percent) + 
    labs(x = xtitle, y = "") +
    theme(axis.text.x = element_text(angle = 40, hjust = 1))
}

#plot graphs together
WillingStudyGraph <- bargraph(data, data$WillingnessStudy, "Length Willing to Study")
LengthStudyGraph <- bargraph(willstudy, willstudy$LengthWillingStudy, "Length Willing to Study")
TimeCommitmentGraph <- bargraph(willstudy, willstudy$WeeklyTimeCommitment, "Weekly Time Commitment")
AttendanceGraph <- bargraph(willstudy, willstudy$AttendanceFrequency, "Frequency of Attendance")

plot_grid(WillingStudyGraph, LengthStudyGraph, TimeCommitmentGraph, AttendanceGraph, nrow = 2)

```


**Drivers have the aptitude to learn new skills**


```{r study completion rates, echo = F}

studiedprev <- willstudy %>% filter(StudiedPreviously == "Yes")

#TODO rename Reasons for not completing levels

StudiedGraph <- bargraph(willstudy, willstudy$StudiedPreviously, "Studied Before?")
CompletedStudyGraph <- bargraph(studiedprev, studiedprev$CompletedStudies,"Completed Studies?")
NotCompletedStudyGraph <- bargraph(studiedprev, studiedprev$ReasonNotCompleting,"Barriers to Completion")

plot_grid(StudiedGraph, CompletedStudyGraph, NotCompletedStudyGraph, nrow = 2)

```


** Drivers are prepared to switch to employee/ freelancer role**

```{r switch position}

bargraph(willstudy, willstudy$SwitchPosition, "Prepared to Switch Position")

```


**Drivers have access to computers and internet for self study**

Count laptop, internet at home, mobile data, none

```{r}

#TODO: correct this

data$laptop = ifelse("Personal computer or laptop" %in% data$ResourceAccess, 1, 0)
data$HomeInternet = ifelse("Internet" %in% data$ResourceAccess, 1, 0)

```

