---
title: "Uber driver survey November 2018"
author: "Michelle Hoogenhout"
date: "14 November 2018"
output: html_document
---

# Purpose 
Design a training pilot

# Hypotheses in designing the pilot:

UBER drivers...
- want higher-value economic opportunities / to learn skills
- are interested in digital careers: Trades, digital, business
- have the aptitude for high-value digital careers
- are prepared to study part-time alongside UBER driving long enough to develop employable skills
- are prepared to switch to being an employee or freelancer rather than UBER driver "entrepreneur" on completion of the programme

```{r setup, include=FALSE}

library(ggplot2)
library(tidyverse)
library(magrittr)
library(psych)
library(cowplot) #for plotting
library(scales) #for plotting
library(readxl) #for excel files

```


```{r functions, include = F}

getpercentage <- function(var){
  x <- length(which(var == 1))/length(var)*100
  print(round(x,0))
}

bargraph <- function(data, x, xtitle){
  p <- ggplot(data, aes(x = x))
  p + geom_bar(aes(y = (..count..)/sum(..count..))) + 
    scale_y_continuous(labels=percent) + 
    labs(x = xtitle, y = "") +
    theme(axis.text.x = element_text(angle = 40, hjust = 1))
}

```


```{r define factor levels}

#define order of factor levels
OrderLengthStudy<- c("Less than 6 months", "6 months - 1 year", "1 - 2 years",
                    "3 - 5 years", "More than 5 years")

OrderTimeCommitment<- c("1 - 2 hours", "3 -5 hours", "6 - 10 hours",
                   "10 - 15 hours", "16 - 20 hours", "More than 20 hours")

OrderFreqAttendance<- c("Every weekday", "Twice a week", "Once a week", "Once a month")


#combine all variance of currntly studying
studying_list = c("Currently busy with studying", 
                  "currently studying", 
                  "Currently studying", 
                  "Studies in progress",
                   "I am currently busy studying at college")

```

```{r get data, include=FALSE}

data <- readxl::read_xlsx("Umuzi Driver Education Revised.xlsx")

#rename and create factors, defining order and collapsing levels of currently studying
data %<>% rename(
  id = "Response ID", 
  WillingnessStudy = "Are you willing to study part-time while operating on Uber?",
  LengthWillingStudy = "For how long would you be willing to study while operating on Uber?",
  Skills = "Which skills would you most like to learn? Select up to three skills.",
  DesiredOutcome = "What is most important to you at the end of studying?",
  WeeklyTimeCommitment = "How much time can you commit on a weekly basis to studying?",
  AttendanceFrequency = "How often would you be able to attend classes, in-person, at a local campus in your city?",
  PreferredClassFormat = "What format would you prefer for the classes?",
  ResourceAccess = "I have access to",
  InternetAccess = "How do you access the internet when you not operating on Uber?",
  PreferredCost = "How much would you be willing to pay monthly for your part-time studying?",
  SwitchPosition = "If you were given the opportunity to start a new career, would you be willing to start in a junior position?",
  StudiedPreviously = "Have you studied before?",
  StudySubject = "If \" yes\", what did you study?",
  CompletedStudies = "Did you complete your studies?",
  ReasonNotCompleting = "What prevented you from completing your studies?",
  FocusGroup = "Would you be willing to attend a focus group to talk about your goals for further education and skills?") %>% 
  mutate_all(funs(factor)) %>%
  mutate(LengthWillingStudy = factor(LengthWillingStudy, levels = OrderLengthStudy),
         WeeklyTimeCommitment = factor(WeeklyTimeCommitment, levels = OrderTimeCommitment),
         AttendanceFrequency = factor(AttendanceFrequency, levels = OrderFreqAttendance),
         ReasonNotCompleting = fct_collapse(ReasonNotCompleting, `Currently Studying` = studying_list)
                )

#get laptop and personal computer separately from list
data %<>% mutate(
  laptop=ifelse(grepl("Personal computer", ResourceAccess), 1, 0),
  HomeInternet=ifelse(grepl("Internet", ResourceAccess), 1, 0),
  FriendsPC=ifelse(grepl("friend or family member's computer", ResourceAccess), 1, 0) )

#change Studied Previously frm yes to Currently Studying
levels(data$StudiedPreviously) <- c("No", "Yes", "Currently Studying")
data[which(data$ReasonNotCompleting == "Currently Studying"),]$StudiedPreviously <- "Currently Studying"


#trim whitespace
trimws(data$Skills)

```
## Drivers want higher-value economic opportunities / to learn skills

What percentage of drivers want to study?
Mode and distibution of willingness to spend
Count of certificate, new skill, job

```{r cost outcome willing}

willstudy <- data[which(data$WillingnessStudy == "Yes"),]
 

WillingStudyGraph <- bargraph(data, data$WillingnessStudy, "Length Willing to Study")

bargraph(willstudy, willstudy$PreferredCost, "Preferred Cost")

bargraph(willstudy, willstudy$DesiredOutcome, "")


```


## What skills do drivers want to learn?
rank skills offered; group by trade, business, digital, other

```{r skill types}
# 
# temp <- unlist(strsplit(as.character(data$Skills), c(",", " and ")))
# skillslist <- unlist(strsplit(temp, ","))         
#          
# TRADESKILLS = c("Electrical", "Mechanics", "Construction", "Plumbing", "Rigging", 
#                 " Crane operator", "Boiler Maker", "Boiler making", "Oil and gas drilling operations and reading operations")
# 
# BUSINESS_SKILLS = c("Entrepreneurship (running your own business)", 
#                     "Business Studies", "Business strategy", "Project Management", 
#                     "Financial Accounting", "Accounting", "MBA")
# 
# IT_SKILLS = c("Basic computer literacy", "IT", "IT fixing computers, 
#               Web development", "Data science", "Data engineering", "Programming", "Data analysis")
# 
# MARKETING = "Marketing"
# 
# 
# head(ifelse(TRADESKILLS %in% data$Skills, 1, 0))
# head(data$Skills)
# 
# head(ifelse(BUSINESS_SKILLS %in% data$Skills, 1, 0))
# 
# #add categories related to business, trade and IT
# data %>% mutate(
#   SkillType = Skills,
#   fct_collapse(SkillType, 
#                Trades = TRADESKILLS, 
#                Business = BUSINESS_SKILLS, 
#                Computing = IT_SKILLS, 
#                Marketing = MARKETING)) %>% 
#   add_count(SkillType)
# 
# 
# sort(table(skillslist))
# 
# 

```

## Drivers are prepared to study for long enough to develop skills

```{r willingness to learn, echo = F}

LengthStudyGraph <- bargraph(willstudy, willstudy$LengthWillingStudy, "Length Willing to Study")
TimeCommitmentGraph <- bargraph(willstudy, willstudy$WeeklyTimeCommitment, "Weekly Time Commitment")
AttendanceGraph <- bargraph(willstudy, willstudy$AttendanceFrequency, "Frequency of Attendance")

plot_grid(WillingStudyGraph, LengthStudyGraph, TimeCommitmentGraph, AttendanceGraph, nrow = 2)

```


## Drivers have the aptitude to learn new skills


```{r study completion rates, echo = F}

studiedprev <- willstudy %>% filter(StudiedPreviously == "Yes")

#TODO rename Reasons for not completing levels

barriers = c("Educational support", "Money", "Time", "Started Working", "Prefer not to say", "Personal")

StudiedGraph <- bargraph(willstudy, willstudy$StudiedPreviously, "Studied Before?")
CompletedStudyGraph <- bargraph(studiedprev, studiedprev$CompletedStudies,"Completed Studies?")
NotCompletedStudyGraph <- bargraph(studiedprev, studiedprev$ReasonNotCompleting,"Barriers to Completion")

plot_grid(StudiedGraph, CompletedStudyGraph, NotCompletedStudyGraph, nrow = 2)

```


## Drivers are prepared to switch to employee/ freelancer role

```{r switch position}

bargraph(willstudy, willstudy$SwitchPosition, "Prepared to Switch Position")

```


## Drivers have access to computers and internet for self study

Of the drivers surveyed, `r getpercentage(data$laptop)` have laptops, `r getpercentage(data$HomeInternet)` have internet at home. Most drivers access the internet through their mobile phone. 

```{r laptop internet}

table(data$laptop)
table(data$HomeInternet)
table(data$InternetAccess)


getpercentage(data$laptop)
getpercentage(data$HomeInternet)
getpercentage(data$FriendsPC)


bargraph(willstudy, willstudy$InternetAccess, "Method of Internet Access")

```

