---
title: "Uber driver survey November 2018"
author: "Michelle Hoogenhout"
date: "14 November 2018"
output: html_document
---

# Purpose 
Design a training pilot

# Hypotheses in designing the pilot:

UBER drivers...
- want higher-value economic opportunities / to learn skills
- are interested in digital careers: Trades, digital, business
- have the aptitude for high-value digital careers
- are prepared to study part-time alongside UBER driving long enough to develop employable skills
- are prepared to switch to being an employee or freelancer rather than UBER driver "entrepreneur" on completion of the programme

```{r setup, include=FALSE}

library(tidyverse) #data wrangling
library(magrittr) #pipe
library(ggplot2) #plot
library(scales) #for plotting
library(readxl) #for excel files
library(tm) #word count 
library(tau)

```


```{r functions, include = F}

#get percentage of TRUE observations
getpercentage <- function(var){
  x <- length(which(var == 1))/length(var)*100
  print(round(x,0))
}

#basic bargraph
bargraph <- function(data, x, title = ""){
  p <- ggplot(data, aes(x = x))
  p + geom_bar(aes(y = (..count..)/sum(..count..))) + 
    scale_y_continuous(labels=percent) + 
    labs(x = "", y = "") +
    ggtitle(title) +
    theme(axis.text.x = element_text(angle = 40, size = 12, hjust = 1),
          axis.text.y = element_text(size = 12))
    
}

#export graph
exportgraph <- function(graphname, exportname, width=8, height=6){
  png(filename = exportname, 
    type="cairo",
    units="in", 
    width=width, 
    height=height, 
    pointsize=12, 
    res=96)
  return(graphname)
dev.off()

}

#count number of observations fitting within defined category in a list
countcats <- function(category, list){
  length(grep(paste(category,collapse="|"), 
                        list, value=TRUE))
}


#get count of words that occur frequently
wordcount <- function(minfreq, maxfreq, list, removewords){

  # Remove Words and Tokenize Text
  skillstext <- textcnt(
  removeWords(scan_tokenizer(list), removewords), 
  method = "string", n = 1L, lower = 1L)

  # Change List to Data Frame
  skillsdata <- plyr::ldply(skillstext, data.frame) 

  # Filter by occurence
  results <- skillsdata %>%  filter(skillsdata[,2]> minfreq & skillsdata[,2]< maxfreq)
  colnames(results)<-c("word", "frequency")
  
  return(results)
  
}

```


```{r define factor levels}

#define order of factor levels
OrderLengthStudy<- c("Less than 6 months", "6 months - 1 year", "1 - 2 years",
                    "3 - 5 years", "More than 5 years")

OrderTimeCommitment<- c("1 - 2 hours", "3 -5 hours", "6 - 10 hours",
                   "10 - 15 hours", "16 - 20 hours", "More than 20 hours")

OrderFreqAttendance<- c("Every weekday", "Twice a week", "Once a week", "Once a month")

OrderPreferredCost <- c("I'm not willing to pay for studying", "Less than R200", 
                        "R200 - R499", "R500 - R999", "R1000 - R1999", "R2000 - R2999")


#combine all variance of currntly studying
studying_list = c("Currently busy with studying", 
                  "currently studying", 
                  "Currently studying", 
                  "Studies in progress",
                   "I am currently busy studying at college")

```

```{r get data, include=FALSE}

data <- readxl::read_xlsx("Umuzi Driver Education Revised.xlsx")

#rename and create factors, defining order and collapsing levels of currently studying
data %<>% rename(
  id = "Response ID", 
  WillingnessStudy = "Are you willing to study part-time while operating on Uber?",
  LengthWillingStudy = "For how long would you be willing to study while operating on Uber?",
  Skills = "Which skills would you most like to learn? Select up to three skills.",
  DesiredOutcome = "What is most important to you at the end of studying?",
  WeeklyTimeCommitment = "How much time can you commit on a weekly basis to studying?",
  AttendanceFrequency = "How often would you be able to attend classes, in-person, at a local campus in your city?",
  PreferredClassFormat = "What format would you prefer for the classes?",
  ResourceAccess = "I have access to",
  InternetAccess = "How do you access the internet when you not operating on Uber?",
  PreferredCost = "How much would you be willing to pay monthly for your part-time studying?",
  SwitchPosition = "If you were given the opportunity to start a new career, would you be willing to start in a junior position?",
  StudiedPreviously = "Have you studied before?",
  StudySubject = "If \" yes\", what did you study?",
  CompletedStudies = "Did you complete your studies?",
  ReasonNotCompleting = "What prevented you from completing your studies?",
  FocusGroup = "Would you be willing to attend a focus group to talk about your goals for further education and skills?") %>% 
  mutate_all(funs(factor)) %>%
  mutate(LengthWillingStudy = factor(LengthWillingStudy, levels = OrderLengthStudy),
         WeeklyTimeCommitment = factor(WeeklyTimeCommitment, levels = OrderTimeCommitment),
         AttendanceFrequency = factor(AttendanceFrequency, levels = OrderFreqAttendance),
         PreferredCost = factor(PreferredCost, levels = OrderPreferredCost),
         ReasonNotCompleting = fct_collapse(ReasonNotCompleting, `Currently Studying` = studying_list)
                )

#get laptop and personal computer separately from list
data %<>% mutate(
  laptop=ifelse(grepl("Personal computer", ResourceAccess), 1, 0),
  HomeInternet=ifelse(grepl("Internet", ResourceAccess), 1, 0),
  FriendsPC=ifelse(grepl("friend or family member's computer", ResourceAccess), 1, 0) )

#change Studied Previously frm yes to Currently Studying
levels(data$StudiedPreviously) <- c("No", "Yes", "Currently Studying")
data[which(data$ReasonNotCompleting == "Currently Studying"),]$StudiedPreviously <- "Currently Studying"

#trim whitespace and make lower case for differences in input
data$skills <- tolower(trimws(data$Skills))

```
## Drivers want higher-value economic opportunities / to learn skills

What percentage of drivers want to study?
Mode and distibution of willingness to spend
Count of certificate, new skill, job

```{r cost outcome willing}

willstudy <- data[which(data$WillingnessStudy == "Yes"),]
 

WillingStudyGraph <- bargraph(data, data$WillingnessStudy, "Length Willing to Study")

costgraph <- bargraph(willstudy, willstudy$PreferredCost, "Preferred Cost")

outcomegraph <- bargraph(willstudy, willstudy$DesiredOutcome, "Desired Outcome")

```


## What skills do drivers want to learn?
rank skills offered; group by trade, business, digital, other

```{r skill types}

#List all Skills in vector
temp <- unlist(strsplit(as.character(data$Skills), c(",", " and ")))
skillslist <- unlist(strsplit(temp, ","))

#categories
TRADESKILLS = c("Electrical", "Mechanics", "Construction", "Plumbing", "Rigging",
                " Crane operator", "Boiler Maker", "Boiler making", 
                "Oil and gas drilling operations and reading operations")

BUSINESS_SKILLS = c("Entrepreneurship (running your own business)",
                    "Business Studies", "Business strategy", "Project Management",
                    "Financial Accounting", "Accounting", "MBA")

IT_SKILLS = c("Basic computer literacy", "IT", "IT fixing computers,
              Web development", "Data science", "Data engineering", "Programming", 
              "Data analysis")

MARKETING = "Marketing"

skillcats <- list(TRADESKILLS, BUSINESS_SKILLS, IT_SKILLS, MARKETING)


#count skills by category
skillscount <- list()

for (cat in seq_along(skillcats)){
  skillscount[[cat]] <- countcats(skillcats[[cat]], skillslist)
}

#Bind to df
count <- unlist(skillscount)
skill_df <- data.frame(c("Trade", "Business", "IT", "Marketing"), count)
colnames(skill_df) <- c("Category", "Count")

#graph popular skill areas 
SkillGraph <- ggplot(skill_df, aes(Category, Count)) +
         geom_bar(stat = "identity") + 
         ggtitle

```

```{r}
#Visualise all reported skills

# words not to be counted
removewords <- c("and", "basic", "literacy", "making", 
                "development", "running", "your", "own", "the", " basic")

skillcount <- wordcount(3,100, skillslist, removewords)

#polar coordinates graph of all words
SkillCoord <- ggplot(skillcount, aes(reorder(word, frequency), frequency, fill=word)) + 
  geom_bar(stat = "identity", colour = "black") +
  coord_polar(theta = "x") + xlab("") + ylab("") + 
  ggtitle("Word Frequency: Skills") +
  theme(legend.position = "none",
        axis.text.x = element_text(size = 12, angle=-10)) + 
  labs(x = NULL, y = NULL)

#bargraph
SkillBarGraph <- ggplot(skillcount, aes(reorder(word, frequency), frequency)) +
  geom_bar(stat = "identity", colour = "black") + 
  ggtitle("Word Frequency: Skills") + labs(y = "Frequency", x = "") +
  theme(axis.text.x = element_text(size = 12, angle = 40, hjust = 1))

#colourful
plotly::ggplotly(ggplot2::ggplot(skillcount, aes(reorder(word, frequency), frequency, fill=word)) +
                   geom_bar(width = 0.75, stat = "identity", colour = "black", size = 1) + 
                   xlab("") + ylab("") + 
                   ggtitle("Word Frequency") + 
                   theme(legend.position = "none") + 
                   labs(x = NULL, y = NULL) + 
                   theme(plot.subtitle = element_text(vjust = 1), 
                         plot.caption = element_text(vjust = 1), 
                         axis.text.x = element_text(size = 12, angle = 40)) + 
                   theme(panel.background = element_rect(fill = "honeydew1"),
                         plot.background = element_rect(fill = "antiquewhite"))) 

```

## Drivers are prepared to study for long enough to develop skills

```{r willingness to learn, echo = F}

LengthStudyGraph <- bargraph(willstudy, willstudy$LengthWillingStudy, "Length Willing to Study")
TimeCommitmentGraph <- bargraph(willstudy, willstudy$WeeklyTimeCommitment, "Weekly Time Commitment")
AttendanceGraph <- bargraph(willstudy, willstudy$AttendanceFrequency, "Frequency of Attendance")

```


## Drivers have the aptitude to learn new skills


```{r study completion rates, echo = F}

studiedprev <- willstudy %>% filter(StudiedPreviously == "Yes")

#TODO rename Reasons for not completing levels
barriers = c("Educational support", "Money", "Time", "Started Working", "Prefer not to say", "Personal")

StudiedGraph <- bargraph(willstudy, willstudy$StudiedPreviously, "Studied Before?")
CompletedStudyGraph <- bargraph(studiedprev, studiedprev$CompletedStudies,"Completed Studies?")

NotCompletedStudyGraph <- ggplot(
  studiedprev, aes(ReasonNotCompleting)) +
  geom_bar(aes(y = (..count..)/sum(..count..))) + 
  ggtitle("Barriers to Completion") +
  scale_x_discrete(labels = barriers) +
  theme(axis.text.x = element_text(angle=40, size = 12, hjust = 1),
        axis.text.y = element_text(size = 12)) +
  labs(x = "", y = "")
  
```


What did drivers study before?
```{r}


#polar coordinates graph of all words
PrevStudiesCoord <- ggplot(studiedprev, aes(StudySubject, fill=StudySubject)) + 
  geom_bar(colour = "black") +
  coord_polar(theta = "x") + xlab("") + ylab("") + 
  theme(legend.position = "none",
        axis.text.x = element_text(size = 6, angle=-10)) + 
  labs(x = NULL, y = NULL)

#bargraph
PrevStudiesGraph <- ggplot(studiedprev, aes(StudySubject, fill=StudySubject)) +
  geom_bar(colour = "black") + 
  labs(y = "Frequency", x = "") +
  theme(axis.text.x = element_text(size = 12, angle = 40, hjust = 1))

```


## Drivers are prepared to switch to employee/ freelancer role

```{r switch position}

SwitchPositionGraph <- bargraph(willstudy, willstudy$SwitchPosition, "Prepared to Switch Position")

```


## Drivers have access to computers and internet for self study

Of the drivers surveyed, `r getpercentage(data$laptop)` have laptops, `r getpercentage(data$HomeInternet)` have internet at home. Most drivers access the internet through their mobile phone. 


```{r laptop internet}

table(data$laptop)
table(data$HomeInternet)
table(data$InternetAccess)


getpercentage(data$laptop)
getpercentage(data$HomeInternet)
getpercentage(data$FriendsPC)


InternetGraph <- bargraph(willstudy, willstudy$InternetAccess, "Method of Internet Access")

```

```{r graph export}

graphlist <- list(c(outcomegraph, "figs/Outcome.png"),
              c(costgraph, "figs/Cost.png"),
              c(LengthStudyGraph, "figs/LengthStudy.png"),
              c(LengthStudyGraph, "figs/LengthStudy.png"),
              c(TimeCommitmentGraph, "figs/TimeCommitment.png"),
              c(AttendanceGraph, "figs/Attendance.png"),
              c(StudiedGraph, "figs/StudiedBefore.png"),
              c(CompletedStudyGraph, "figs/CompletedStudies.png"),
              c(NotCompletedStudyGraph, "figs/ReasonsNotCompletedStudy.png"),
              c(SwitchPositionGraph, "figs/SwitchPosition.png"),
              c(InternetGraph, "figs/InternetAccess.png"))
# 
# for (graph in seq_along(graphlist)){
#   exportgraph(bquote(.(graphlist[graph][1])), bquote(.(graphlist[graph[2]])))
# }


exportgraph(outcomegraph, "figs/Outcome.png")
exportgraph(costgraph, "figs/Cost.png")
exportgraph(LengthStudyGraph, "figs/LengthStudy.png")
exportgraph(TimeCommitmentGraph, "figs/TimeCommitment.png")
exportgraph(AttendanceGraph, "figs/Attendance.png")
exportgraph(StudiedGraph, "figs/StudiedBefore.png")
exportgraph(CompletedStudyGraph, "figs/CompletedStudies.png")
exportgraph(NotCompletedStudyGraph, "figs/ReasonsNotCompletedStudy.png")
exportgraph(SwitchPositionGraph, "figs/SwitchPosition.png")
exportgraph(InternetGraph, "figs/InternetAccess.png")
exportgraph(SkillGraph, "figs/SkillCategories.png")
exportgraph(SkillCoord, "figs/SkillCategories_polar.png")
exportgraph(SkillBarGraph, "figs/SkillCategoriesAll.png")
exportgraph(PrevStudiesCoord, "figs/PrevStudySubject_coord.png")
exportgraph(PrevStudiesGraph, "figs/PrevStudySubject.png")


dev.off()
```

